---
title: "FML Assignment_2"
author: "Sri Hari Vadhry"
date: "2023-10-08"
output:
  word_document: default
  html_document: default
---

# Summary

## Questions - Answers

1. How would this customer be classified?
A. This new customer would be classified as 0, does not take the personal loan


2. What is a choice of k that balances between overfitting and ignoring the predictor
information?
A. The best of K is 3 because of the efficiency.


3. Show the confusion matrix for the validation data that results from using the best k.
A.  By using K as 3, the confusion matrix is
          Reference
Prediction    0    1
         0 1786   63
         1    9  142

True Positive- 142
True Negative- 1786
False Positive- 9
False Negative- 63

4. Consider the following customer: Age = 40, Experience = 10, Income = 84,Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit Card = 1. Classify the customer using the best k.

A. By using K= 3, the customer classified to be 0, So he does not take Personal Loan.

5. Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply the k-NN method with the k chosen above. Comparision of the confusion matrix of the test set with that of the training and validation sets.
A.

For Training the Accuracy is 97.64%, Sensitivity is 99.78% and Specificity is 76.72%
For Validation the Accuracy is 96.8%, Sensitivity is 99.56% and Specificity is 69.12%
For Test the Accuracy is 96.1%, Sensitivity is 99.55% and Specificity is 68.75%

We can observe that the accuracy, sensitivity and specificity is comparatively high for training data because of overfitting of training data

#Test v/s Training:
Accuracy:We can see that training dataset has higher accuracy compared to test, as training dataset may be more balanced dataset or easier to predict.

Sensitivity(True positive rate):Training dataset has higher sensitivity rate upon comparison with test dataset, there by indicating that training model is more better at correctly identifying positive cases.

Specificity(True Negative rate):Training dataset has higher specificity as when compared to test dataset, there by indicating that training model is more better at correctly identifying negative cases.
.

#Test v/s Validation:
Accuracy:We can see that validation dataset has higher accuracy compared to test, as validation dataset may be more balanced dataset or easier to predict.

Sensitivity(True positive rate):Validation dataset has slightly higher sensitivity rate upon comparison with test dataset, there by indicating that validation model is more better at correctly identifying positive cases.

Specificity(True Negative rate):Validation dataset has higher specificity as when compared to test dataset, there by indicating that training model is more better at correctly identifying negative cases.



# Problem Statement

Universal bank is a young bank growing rapidly in terms of overall customer acquisition.
The majority of these customers are liability customers (depositors) with varying sizes of relationship with the bank. The customer base of asset customers (borrowers) is quite
small, and the bank is interested in expanding this base rapidly in more loan business. In particular, it wants to explore ways of converting its liability customers to personal loan customers.

A campaign that the bank ran last year for liability customers showed a healthy conversion rate of over 9% success. This has encouraged the retail marketing department to devise smarter campaigns with better target marketing. The goal is to use k-NN to predict whether a new customer will accept a loan offer. This will serve as the basis for the design of a new campaign.

The file UniversalBank.csv contains data on 5000 customers. The data include customer
demographic information (age, income, etc.), the customer’s relationship with the bank
(mortgage, securities account, etc.), and the customer response to the last personal loan
campaign (Personal Loan). Among these 5000 customers, only 480 (= 9.6%) accepted the
personal loan that was offered to them in the earlier campaign.

Partition the data into training (60%) and validation (40%) sets

# Installing the required packages

```{r}

# install.packages("caret")
# install.packages("e1071")

```

# Loading the required libraries

```{r}

library(class)
library(caret)
library(e1071)

```

# Data Importing

```{r}
universal_bank <- read.csv("C:/Users/Hari/OneDrive/Desktop/FML/UniversalBank.csv")

dim(universal_bank)

t(t(names(universal_bank)))


```

# Data Cleaning

#Drop ID and ZIP.Code
```{r}

universal_bank <- universal_bank[,-c(1,5)]

```

#Education needs to be converted to factor

```{r}

universal_bank$Education <- as.factor(universal_bank$Education)  

#Convert Education to dummy variables
groups <- dummyVars(~., data = universal_bank)

universal_updated <- as.data.frame(predict(groups,universal_bank))

```

# Split the data into 60% training and 40% validation

```{r}
# Important to ensure that we get the same sample if we rerun the code

set.seed(1)  

train.index <- sample(row.names(universal_updated), 0.6*dim(universal_updated)[1])

valid.index <- setdiff(row.names(universal_updated), train.index)  

train.df <- universal_updated[train.index,]

valid.df <- universal_updated[valid.index,]

t(t(names(train.df)))

```

# Normalising the Data

```{r}
train.norm.df <- train.df[,-10] # Note that Personal Income is the 10th variable
valid.norm.df <- valid.df[,-10]

norm.values <- preProcess(train.df[, -10], method=c("center", "scale"))

train.norm.df <- predict(norm.values, train.df[, -10])
valid.norm.df <- predict(norm.values, valid.df[, -10])

```

1. Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code using k = 1. Remember to transform categorical predictors with more than two categories into dummy variables first. Specify the success class as 1 (loan acceptance), and use the default cutoff value of 0.5. How would this customer be classified?



```{r}

new_customer <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)

# Normalize the new customer
new.cust.norm <- new_customer
new.cust.norm <- predict(norm.values, new.cust.norm)


```


Now, let us predict using knn
```{r}

knn.pred1 <- class::knn(train = train.norm.df, 
                       test = new.cust.norm, 
                       cl = train.df$Personal.Loan, k = 1)
knn.pred1

```
2. What is a choice of k that balances between overfitting and ignoring the predictor
information?

```{r}
# Calculate the accuracy for each value of k
# Set the range of k values to consider

accuracy.df <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))
for(i in 1:15) {
  knn.pred <- class::knn(train = train.norm.df, 
                         test = valid.norm.df, 
                         cl = train.df$Personal.Loan, k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.pred, 
                                       as.factor(valid.df$Personal.Loan),positive = "1")$overall[1]
}

which(accuracy.df[,2] == max(accuracy.df[,2])) 

plot(accuracy.df$k,accuracy.df$overallaccuracy, main = "Accuracy vs Overall Accuracy", col= "blue")



```

3. Show the confusion matrix for the validation data that results from using the best k.

```{r}

knn.pred2 <- class::knn(train = train.norm.df,
                        test = valid.norm.df,
                        cl= train.df$Personal.Loan, k= 3)
knn.pred2

confusion.matrix <- confusionMatrix(knn.pred2, as.factor(valid.df$Personal.Loan), positive = "1")
confusion.matrix


```

4. Consider the following customer: Age = 40, Experience = 10, Income = 84,
Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0,
Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit
Card = 1. Classify the customer using the best k.

```{r}

new_customer1 <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)

# Normalize the new customer
new.cust.norm1 <- new_customer1
new.cust.norm1 <- predict(norm.values, new.cust.norm1)

# knn prediction

knn.pred3 <- class::knn(train = train.norm.df,
                        test = new.cust.norm1,
                        cl= train.df$Personal.Loan, k= 3)
knn.pred3


```
5. Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply the k-NN method with the k chosen above. Comparision of the confusion matrix of the test set with that of the training and validation sets.

```{r}
#Splitting the data into training set(50%), validation set(30%) and testing set(20%)

set.seed(1)  

train.index1 <- sample(row.names(universal_updated), 0.5*dim(universal_updated)[1])
train.df1 <-universal_updated[train.index1,]

valid.index1 <- setdiff(row.names(universal_updated), train.index1)
valid.df1 <- universal_updated[valid.index1, ]

valid.index2 <- sample(row.names(valid.df1), 0.6*dim(valid.df1)[1])
valid.df2 <- valid.df1[valid.index2, ]

test.index1 <- setdiff(row.names(valid.df1),valid.index2)
test.df1 <- valid.df1[test.index1, ]


```

```{r}
#Normalize the above data
train.norm.df1 <- train.df1[,-10]
valid.norm.df2 <- valid.df2[,-10]
test.norm.df1 <- test.df1[,-10]

norm.values1 <- preProcess(train.df1[,-10], method = c("center", "scale"))

train.norm.df1 <- predict(norm.values1, train.df1[,-10])
valid.norm.df2 <- predict(norm.values1, valid.df2[,-10])

test.norm.df1 <- predict(norm.values1, test.df1[,-10])
```

#Knn-prediction (training data - 50%)
```{r}

knn.pred4 <- class::knn(train = train.norm.df1,
                        test = train.norm.df1,
                        cl= train.df1$Personal.Loan, k= 3)
knn.pred4

confusion.matrix1 <- confusionMatrix(knn.pred4, as.factor(train.df1$Personal.Loan))
confusion.matrix1

```
#Knn-prediction (Validation data - 30%)
```{r}
knn.pred5 <- class::knn(train = train.norm.df1,
                        test = valid.norm.df2,
                        cl= train.df1$Personal.Loan, k= 3)
knn.pred5

confusion.matrix2 <- confusionMatrix(knn.pred5, as.factor(valid.df2$Personal.Loan))
confusion.matrix2

```

#Knn-prediction (Test data - 20%)
```{r}
knn.pred6 <- class::knn(train = train.norm.df1,
                        test = test.norm.df1,
                        cl= train.df1$Personal.Loan, k= 3)
knn.pred6

confusion.matrix3 <- confusionMatrix(knn.pred6, as.factor(test.df1$Personal.Loan))
confusion.matrix3
```

