---
title: "FML Assignment 4"
output:
  word_document: default
  html_document: default
date: "2023-11-11"
---

```{r}

library(ggplot2)
library(factoextra)
library(flexclust)
library(cluster)
library(tidyverse)
library(dplyr)

```

#   1.Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on.

```{r}

ps <- read.csv("C:/Users/sriha/Desktop/Assignments/Fundamentals of Machine Learning/Datasets/Pharmaceuticals.csv")

pharmaceuticals <- na.omit(ps)

pharmaceuticals

```


# Considering Quantitative data
```{r}
row.names(pharmaceuticals)<- pharmaceuticals[,1]

pharma <- pharmaceuticals[,3:11]

t(t(names(pharma)))

head(pharma)

```

# Normalising the Data

```{r}

pharma1<-scale(pharma)

head(pharma1)


```

# Elbow method for determining number of clusters

```{r}

fviz_nbclust(pharma1, kmeans, method = "wss")

# From the graph we cannot determine K value, Hence using silhoutte method

```
#Silhouette method for determining number of clusters

```{r}

fviz_nbclust(pharma1, kmeans, method = "silhouette")



```


# K means clustrering

```{r}
set.seed(150)

k5 <- kmeans(pharma1, centers = 5, nstart = 25)

#Visualize the output

k5$centers  #centroids

```


```{r}
#  Visualizing the clusters

fviz_cluster(k5,data = pharma1) 


```
#Cluster 1 - AGN, PHA (low Asset_Turnover, High PE Ratio)

#Cluster 2 - WPI, MRX,ELN,AVE (low PE_Ratio,high ROE,low ROA,low Net_Profit_Margin, high Rev_Growth)

#Cluster 3 - AHM,SGP,WYE,BMY,AZN, ABT, NVS, LLY ( low Market_Cap,low Beta,low PE_Ratio,high Leverage,high Rev_Growth.)

#Cluster 4 - JNJ, MRK, PFE,GSK (High Market_Cap,ROE, ROA,Asset_Turnover Ratio and low Beta/PE Ratio)

#Cluster 5 - BAY, CHTT, IVX (low Rev_Growth,high Beta and levearge,low Net_Profit_Margin)



```{r}

print(k5)

```

```{r}

dis <- dist(pharma1, method = "euclidean")

fviz_dist(dis)


```


```{r}

# From thhe data the center is defined after 25 restarts and there are 5 clusters. 

fit <- kmeans(pharma1,5)

```

```{r}

#Finding the mean value of variables

aggregate(pharma1,by=list(fit$cluster),FUN=mean)


```

```{r}

# Cluster Plot

clusplot(pharma1,fit$cluster,color =
           TRUE,shade = TRUE,labels = 2,lines = 0)


```

## 2 Interpret the clusters with respect to the numerical variables used in forming the clusters. Is there a pattern in the clusters with respect to the numerical variables?

```{r}

clust_data <- ps[12:14]
clust_data$Clusters <- k5$cluster
ggplot(clust_data, aes(factor(Clusters), fill= Median_Recommendation))+ geom_bar(position='dodge')+labs(x='Clusters')


```
# From the above graph the cluster 1 is having low profit and it has moderate leverage. The graph determines same level of moderate buy and Hold.
# In cluster 2, it is having high revenue growth there might be a potential growth in these companies because it has equal level of moderate buy and sell
# In cluster 3, There is high level of Hold and also the strong large amount of profit which will attract the investors to buy more of this cluster.
# In cluster 4, There is high amount of market captalization and also high profit. It represents high potential of the company and has equal level of hold and buy.
# In cluster 5, The graph represents high debt that leads to high leverage.


```{r}

ggplot(clust_data, aes(factor(Clusters), fill= Location))+ geom_bar(position='dodge')+labs(x='Clusters')


```

```{r}

ggplot(clust_data, aes(factor(Clusters), fill= Exchange))+ geom_bar(position='dodge')+labs(x='Clusters')


```



```{r}
# 3.Provide an appropriate name for each cluster using any or all of the variables in the dataset.

# Cluster 1: Significant Risk

# Cluster 2: Potential Growth

# Cluster 3: Perfect Asset

# Cluster 4: Long term investment

# Cluster 5: Short term investment


```



